---
title: "lab9_part1"
author: "R. Spellenberg"
date: "2023-03-14"
output: 
  html_document:
    code_folding: hide
---

```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(tsibble) ### time series
library(feasts)
library(fable)
library(broom)
```

## Read in our energy data and convert to a tsibble

Data here are from the US Energy Information Administration. US Residential energy consumption from Jan 1973 - Oct 2017. Units = trillions of BTUs

```{r}
energy_df <- read_csv(here('data/energy.csv'))

energy_ts <- energy_df %>% 
  mutate(date = tsibble::yearmonth(month)) %>% 
  as_tsibble(key = NULL, index = date ) ### index tells what column to look at

### head(energy_ts) in console tells us class of columns
```

## Exploratory times series visualization

```{r}
ggplot(data = energy_ts, aes(x = date, y = res_total)) +
  geom_line() +
  labs(y = 'Residential energy consumption\n (Trillion BTU)') +
  theme_minimal()
```

### Season Plot

```{r}
energy_ts %>% 
  gg_season(y = res_total) + ### feasts package
  theme_minimal() +
  scale_color_viridis_c() +
  labs(x = 'Month', y = 'Residential energy consumption\n (Trillion BTU)')
```

### Subseries plot

```{r}
energy_ts %>% 
  gg_subseries(y = res_total)
```

## Decomposition

STL decomposition vs. classical decomposition - classical has same seasonality over entire time series, but STL allows seasonality to change over time
```{r}
### calculate STL decomposition
dcmp <- energy_ts %>% 
  model(STL(res_total ~ season()))

### components(dcmp) in console

### visualize
components(dcmp) %>% 
  autoplot() + 
  theme_minimal()
```

## Autocorrelation function (ACF)

```{r}
energy_ts %>% 
  ACF(y = res_total) %>%  ###feasts package
  autoplot() +
  theme_minimal()
```

## Forecasting Holt-Winters exponential smoothing
(fits with certain types of seasonality)
exponential smoothing is paying attention more to recent events

```{r}
### create our model
energy_fit <- energy_ts %>% 
  model(ets = ETS(res_total ~ season('M'))) ### M for multiplicative, A for additive

### create forecast based on our model, 10 years into the future
energy_forecast <- energy_fit %>% 
  forecast(h = '10 years') ### how far into the future we're going to predict

energy_forecast %>% 
  autoplot()

```

Now plot the actual res_total values on top of the predicted values

```{r}
### append the predicted values (and residuals) to the original energy data
energy_predicted <- broom::augment(energy_fit)

ggplot(data = energy_predicted) +
  geom_line(aes(x = date, y = res_total), color = 'black') +
  geom_line(aes(x = date, y = .fitted), color = 'red', alpha = 0.7)

### want uncorrelated residuals and centered around 0 (mean of 0)/normally distributed
ggplot(data = energy_predicted, aes(x = .resid)) +
  geom_histogram()

ggplot(data = energy_predicted, aes(x = date, y = .resid)) +
  geom_point()
```


